---
title: "DataCamp: Intro to R Assignment"
author: Sarah Guitart
output: 
  html_notebook:
    toc: TRUE
    toc_float: TRUE
    theme: darkly
---


```{r}
# Create the following code and use it to answer Qs 1-6
# A variable a that contains the text of your first name.
# A variable b1 that contains the number 45.6
# A variable b2 that contains the text “45.6”
# A variable c1 that contains the sequence of integers from 0 to 3
a <- "Sarah Guitart"
b1 <- 45.6
b2 <- "45.6"
c1 <- c(0:3)
```


## Q1: What type of data is contained in the variable a?

#### Variable a is a character string 
```{r}
a
class(a)
```


## Q2: What type of data is contained in the variable b1?

#### Variable b1 is numeric
```{r}
b1
class(b1)
```


## Q3: What type of data is contained in the variable b2?

#### Variable b2 is a character string
```{r}
b2
class(b2)
```


## Q4: Explain what happens when you try to add b1 and b2 and why.

#### When I try to add b1 and b2 together, it results in an error because we are trying to add text to a number.
```{r}
x = b1 + b2
```


## Q5: Are the variables b1 and c1 the same type? Why or why not?

#### The variables are not the same because numeric data objects in R can contain decimals, while integers are strictly positive or negative whole numbers.
```{r}
b1
c1
class(b1)
class(c1)
```


## Q6: Explain what happens when you add b1 and c1. Consider both the number of elements in each variable and the data types.

#### When I add the two variables, it adds b1 to all elements within the vector. This makes them numeric, as they are no longer whole numbers. The new object is a vector of 4 numeric variables. 
```{r}
bc <- b1 + c1
bc
class(bc)
```

```{r}
# Create a vector called v1 that contains a sequence of integers from -2 to 2.
# When you print the contents of v1, it should look like this:
#> [1] -2 -1  0  1  2
#Now, use v1 to create a new vector called v2 whose elements are the elements of v1 multiplied by 3. It should look like this:
#> [1] -6 -3  0  3  6
# Finally, calculate the sum of all the elements in v2.

```


## Q7: Show the R code you used to create v1.
```{r}
v1 <- c(-2:2)
v1
```


## Q8: Show the R code you used to create v2.
```{r}
v2 <- v1 * 3
v2
```


## Q9: Show the R code you used to calculate the sum of elements in v2.
```{r}
sum(v2)
```

```{r}
# Create a vector called vec_4 whose elements are the integers from 1 to 12.
# Create a matrix mat_1 from vec_4 that has three rows and four columns. The values in mat_1 should be sequentially increasing by row.
# For example, the first row of mat_1 should contain the values 1, 2, 3, 4.
# Create a matrix mat_2 from vec_4 that has three rows and four columns. The values in mat_2 should be sequentially increasing by column.
#For example, the first column of mat_1 should contain the values 1, 2, 3.
vec_4 <- c(1:12)
vec_4
```

## Q10: Show the code you used to create mat_1.
```{r}
mat_1 <- matrix(data = vec_4, nrow = 3, ncol = 4, byrow = TRUE)
mat_1
```

## Q11: Show the code you used to create mat_2.
```{r}
mat_2 <- matrix(data = vec_4, nrow = 3, ncol = 4)
mat_2
```


```{r}
# Create a list, named my_list_1 with following three elements:
# first element is numeric: 5.2
# second element is a string: “five point two”
# third element is a vector of all integers from 0 to 5 [how do you do this?]
#Name the elements in my_list_1:
#“two”
#“one”
#“three”
```

## Q12: Show the R code you used to create my_list_1.
```{r}
my_list_1 <- list("two" = 5.2, "one" = "five point two", "three" = c(0:5))
my_list_1
```


## Q13: Show valid R code that selects the third element of the list.
```{r}
my_list_1[[3]]
```


## Q14: Show the R code that selects the list element with the name “one”.
```{r}
my_list_1$one
```


```{r}
my_vec = rep(1:3, 5)
my_bool_vec <- my_vec == 3
data.frame(my_vec, my_bool_vec)
```


## Q15: Show the R code that you used to create my_bool_vec.
```{r}
my_bool_vec <- my_vec == 3
my_bool_vec
```


## Q16: Show the R code that you used to subset my_vec using my_bool_vec.
```{r}
my_vec[(my_bool_vec)]
```