---
title: "Analysis of Environmental Data - Lab Exercise #12"
author: "Sarah Guitart"
output: 
  html_notebook:
    toc: TRUE
    toc_float: TRUE
    theme: darkly 
---

Students I Collaborated With: Ian Eggleston 

### Q1

```{r setup_data_q1}
library(here)
bird <- read.csv(here("data", "bird.sub.csv"))
hab <- read.csv(here("data", "hab.sub.csv"))
birdhab <- merge(bird, hab)
dim(birdhab)
fit_1 <- lm(BRCR ~ ls, data = birdhab)
fit_1_coefs = coefficients(fit_1)
fit_1_summary = summary(fit_1)
sd_obs <- fit_1_summary$sigma
int_obs <- fit_1_coefs[1]
slope_obs <- fit_1_coefs[2]
```

```{r create_functions_q1}
linear_simulator = function(x, y_int, slope, st_dev)
{
  noise = rnorm(length(x), mean = 0, sd = st_dev)
  return(slope * x + y_int + noise)
}
linear_sim_fit = function(x, slope, y_int, st_dev)
{
  y_sim = linear_simulator(
    x = x,
    y_int = y_int,
    slope = slope,
    st_dev = st_dev
  )
    fit_sim = lm(y_sim ~ x)
  return(fit_sim)
}
```

```{r create_plot_q1}
alpha = 0.05
n_sims = 30
p_vals = numeric(n_sims)
sample_sizes = seq(2, 20)
sample_size_powers = numeric(length(sample_sizes))
max_x = max(birdhab$ls)
for(j in 1:length(sample_sizes))
{
  x_vals = seq(0, max_x, length.out = sample_sizes[j])
  
  for(i in 1:n_sims)
  {
    fit_sim = linear_sim_fit(
      x = x_vals,
      y_int = int_obs,
      slope = slope_obs,
      st_dev = sd_obs
    )
    p_vals[i] = summary(fit_sim)$coefficients[2, 'Pr(>|t|)']
  }
  sample_size_powers[j] = sum(p_vals < alpha) / n_sims
}
sim_sample_size = 
  data.frame(
    sample_size = sample_sizes,
    power       = sample_size_powers)
```
*Include your final plot in your lab report.*
```{r buildingplot_q1}
fit_lowess_30 = loess(power ~ sample_size, data = sim_sample_size, span = 0.3)
newdata_sample_size = data.frame(sample_size = seq(2, 20, length.out = 100)) 
plot(
  x = newdata_sample_size$sample_size,
  y = predict(fit_lowess_30, newdata = newdata_sample_size),
  type = "l",
  ylab = "Statistical Power", xlab = "Sample Size", main = "Sample Size/Power, LOWESS: 30%")
points(power ~ sample_size, data = sim_sample_size, col = "red", pch = 16)
legend("bottomright", legend = c("Smoothed", "Original"), col = c("black", "red"), lty = c(1, NA), pch = c(NA, 16))
```
### Q2

```{r create_datafunctions_q2}
dat_dispersal <- read.csv(here("data", "dispersal.csv"))
ricker_fun = function(x, a, b) 
{
  return(a * x * exp(-b * x))
}
```

```{r create_fit_q2}
fit_ricker_nls = nls(
  disp.rate.ftb ~ ricker_fun(dist.class, a, b),
  data = dat_dispersal,
  start = list(b = 0, a = 1))
summary(fit_ricker_nls)
```

*Include your plot in your lab report.*
```{r plot_data_q2}
dist_newdata = data.frame(dist.class = seq(0, 1600, length.out = 1600))  
plot(dat_dispersal$dist.class, dat_dispersal$disp.rate.ftb,
    xlim = c(0, 1500), xlab = "Distance Class", ylab = "Standardized Dispersal Rate", 
    main = "Marbled Salamander First Time Breeders - Ricker Model", pch = 17)
lines(predict(fit_ricker_nls, newdata = dist_newdata), col = "blue", lwd = 2)
curve(ricker_fun(x, .6/300, 1/500), add = TRUE, lty = 2, col = "red", lwd = 2)
legend("topright", legend = c("NLS Fit", "Guess Fit", "Data Points"), lwd = c(2,2,NA), lty = c(1,2,NA), col = c("blue", "red", "black"), pch = c(NA,NA,17))
```

### Q3

*What are the AIC values for each of the 4 models?*
```{r create_data_q3}
dat_bird = read.csv(here("data", "bird.sta.csv"))
dat_habitat = read.csv(here("data", "hab.sta.csv"))
dat_all = merge(dat_bird, dat_habitat)
dat_all$GCKI_pres = dat_all$GCKI > 0
fit_gcki_slope = glm(GCKI_pres ~ slope, data = dat_all, family = binomial)
fit_gcki_ba_tot = glm(GCKI_pres ~ ba.tot, data = dat_all, family = binomial)
fit_gcki_both_additive = glm(GCKI_pres ~ slope + ba.tot, data = dat_all, family = binomial)
fit_gcki_both_interactive = glm(GCKI_pres ~ slope * ba.tot, data = dat_all, family = binomial)
```

```{r AIC_values}
AIC(
  fit_gcki_ba_tot,
  fit_gcki_slope,
  fit_gcki_both_additive,
  fit_gcki_both_interactive)
```

A: The AIC values for the four models are 1369.379 for the basal area model, 1432.615 for the slope model, 1355.951 for the additive model, and 1353.007 for the interactive model. 

### Q4

*Which model would you choose, and why?*

A: I would choose the interactive model because it has the lowest AIC value, at 1353.007. This is because the smaller the AIC value, the better the fit of the model. 

### Q5

*Based on the model coefficient table of your chosen model, describe the direction and significance of the relationship(s) of the predictor variable or variables to the binary response. Make sure your answer is in terms of the ecological context.*
```{r}
summary(fit_gcki_both_interactive)
```

A: From the model coefficient table, we can see that the total basal area and the basal area/slope interaction both have associations with Golden-crowned Kinglet presence/absence, as their p-values are below 0.05. The total basal area has a positive relationship of 0.05, while the interaction of slope and total basal area has a negative relationship of -0.0004. Slope by itself also showed a negative association with kinglet presence/absence, but that relationship is not significant. 

### Q6
```{r create_newdata}
n = 500
slope_newdata = data.frame(
  slope = seq(
    from = min(dat_all$slope, na.rm = T),
    to = max(dat_all$slope, na.rm = T),
    length.out = n
  )
)
n = 500
ba_newdata = data.frame(
  ba.tot = seq(
    from = min(dat_all$ba.tot, na.rm = T),
    to = max(dat_all$ba.tot, na.rm = T),
    length.out = n
  )
)
```

```{r predict}
slope_newdata$gcki_predicted = 
  predict(
    fit_gcki_slope,
    newdata = slope_newdata,
    type = "response"
  )
ba_newdata$gcki_predicted = 
  predict(
    fit_gcki_ba_tot,
    newdata = ba_newdata,
    type = "response"
  )
```

*Include your two single-predictor model plots in your report.*
```{r basic_plots}
par(mfrow = c(2, 1))
plot(
  GCKI_pres ~ slope, data = dat_all, axes = FALSE, ylim = c(-0.2,1.2),
  xlab = "Percent Slope", ylab = "", main = "GCKI Presence/Absence by % Slope",
   pch = 16, col = adjustcolor("red", alpha.f = 0.1))
axis(1, at = seq(0,110, by = 10))
axis(2, at = c(0,1), labels = c("Absent", "Present"), las = 2)
box()
lines(gcki_predicted ~ slope, data = slope_newdata)
plot(
  GCKI_pres ~ ba.tot, data = dat_all, axes = FALSE, ylim = c(-0.2, 1.2),
  xlab = "Basal Area", ylab = "", main = "GCKI Presence/Absence by Basal Area",
  pch = 16, col = adjustcolor("blue", alpha.f = 0.1))
axis(1, at = seq(0,210, by = 50))
axis(2, at = c(0,1), labels = c("Absent", "Present"), las = 2)
box()
lines(gcki_predicted ~ ba.tot, data = ba_newdata)
```
### Q7
```{r background_q7}
n = 50
ba.tot = seq(
  from = min(dat_all$ba.tot, na.rm = T),
  to = max(dat_all$ba.tot, na.rm = T),
  length.out = n)
slope = seq(
  from = min(dat_all$slope, na.rm = T),
  to = max(dat_all$slope, na.rm = T),
  length.out = n)
new_dat_all = expand.grid(
  ba.tot = ba.tot,
  slope = slope)
new_dat_all$pred_add = predict(
  fit_gcki_both_additive,
  newdata = new_dat_all,
  type = "response")
new_dat_all$pred_int = predict(
  fit_gcki_both_interactive,
  newdata = new_dat_all,
  type = "response")
z_gcki_add = matrix(
  new_dat_all$pred_add,
  nrow = length(ba.tot),
  byrow = FALSE)
z_gcki_int = matrix(
  new_dat_all$pred_int,
  nrow = length(ba.tot),
  byrow = FALSE)
```

*Include contour plots (or interactive 3D perspective plots) in your report.*
```{r contour_plots}
par(mfrow = c(1, 2))
contour(
  x = ba.tot, y = slope,
  z = z_gcki_add,
  xlab = "Total Basal Area",
  ylab = "Percent Slope",
  main = "Additive")
contour(
  x = ba.tot,
  y = slope,
  z = z_gcki_int,
  xlab = "Total Basal Area",
  ylab = "Percent Slope",
  main = "Interactive")
```
```{r 3d_plots}
require(rgl)
rgl::persp3d(
  x = ba.tot,
  y = slope,
  z = z_gcki_add,
  col = "darkorchid4",
  xlab = "Basal Area",
  ylab = "Slope",
  zlab = "Pr(present)",
  alpha = 0.4)
rglwidget()
rgl::persp3d(
  x = ba.tot,
  y = slope,
  z = z_gcki_int,
  col = "firebrick4",
  xlab = "Basal Area",
  ylab = "Slope",
  zlab = "Pr(present)",
  alpha = 0.4)
rglwidget()
```
